---
title: "Stats 20, 21F -- Homework 4"
author: "TAKAO OBA -- 205615894"
date: "`r format(Sys.Date(), '%D')`"
output:
  html_document:
    toc: yes            # creates table of contents
    toc_depth: 4        # toc will include headers <= ####
    toc_float: yes      # toc always on left of page
    code_folding: show  # allows hiding of code
---


## Question 1

```{r}
my_t <- function(X){
  longvector <- numeric(0)
  if(!is.matrix(X)){
    dim(X) <- c(1, length(X))
    X
  }else{
    for(i in seq_len(dim(X)[2])){
      asvector <- X[, i]
      longvector <- c(longvector, asvector)
    }
    Y <- matrix(longvector, nrow = dim(X)[2], byrow = TRUE)
    Y
  }
}
t(c(1, 2, 3, 4))
my_t(c(1, 2, 3, 4))

A <- matrix(1:6, nrow = 2)
A
t(A)
my_t(A)

B <- matrix(1:10, nrow = 5)
B
t(B)
my_t(B)
```


## Question 2

### (a)

```{r}
my_cor <- function(x, y, use){
  if(use == "pairwise.complete.obs"){
    for(i in length(x)){
      if(is.na(x[i]) | is.na(y[i])){
        x <- x[-i]
        y <- y[-i]
      }
    }
  }
    num <- sum((x - mean(x)) * (y - mean(y)))
    den1 <- sqrt(sum((x - mean(x))^2))
    den2 <- sqrt(sum((y - mean(y))^2))
    r <- num / (den1 * den2)
    r
  
}




```

### (b)

```{r}
linreg <- function(x, y, use){
  if(use == "pairwise.complete.obs"){
    for(i in length(x)){
      if(is.na(x[i]) | is.na(y[i])){
        x <- x[-i]
        y <- y[-i]
      }
    }
  }
  
    b <- (my_cor(x, y, use) * sd(y)) / sd(x)
    a <- mean(y) - b * mean(x)
    c(a, b)

}
```

### (c)

```{r}
x <- c(61, 62, 63, 64, 66, 68)
y <- c(104, 110, 125, 141, 160, 170)
linreg(x, y, "everything")
```

### (d)

```{r}
linreg_mat <- function(x, y){
  X <- cbind(1, x)
  Xt <- my_t(X)
  
  if(dim(Xt)[2] != dim(X)[1]){
    message("Vector arguments are not the same length!")
  }else{
    mat1 <- solve(Xt %*% X)
    if(dim(mat1)[1] != dim(mat1)[2]){
      message("The product of the transpose of X and the matrix X is not a square matrix!")
    }else{
      mat1 %*% Xt %*%y
    }
  }
}
```

### (e)

```{r}
linreg_mat(x, y)
```
Answer from part c is `linreg(x, y, "everything")`.
Answer from part e is `linreg_mat(x, y)`.

### (f)
The value -510.64706 is how much we expect the weight in pounds of the self-identified women to be when their height is 0 inches. 

The value 10.08824 is how much we expect to see an increase in the weight of the self-identified women in pounds when there is one inch increase in their height. 

## Question 3

### (a)

```{r}
`%m%` <- function(A, B){
  row1 <- dim(A)[1]
  row2 <- dim(B)[2]
  C <- matrix(0, nrow = dim(A)[1], ncol = dim(A)[2])
  
  if(dim(A)[2] != dim(B)[1]){
    message("Matrix Multiplication cannot be applied to these two vectors!")
    
  }else{
    for(i in seq_len(dim(A)[1])){
      asvectorA <- A[i, ]
      
      for(j in seq_len(dim(B)[2])){
        
        asvectorB <- B[, j]
        C[i, j] <- sum(asvectorA * asvectorB)
        
      }
    }
    C
  }
}

```

### (b)

```{r}
X <- matrix(c(6, 3, 5, 2, 4, 1), nrow = 2)
Y <- matrix(c(1, 2, 3, 4, 5, 6, 7, 8, 9), nrow = 3)

X %m% Y
```


## Question 4

### (a)

```{r}
`%^%` <- function (A, k) {
  
  if (k == 0){
    B <- diag(dim(A)[1])
  }else if(k == 1){
    B <- A
  }
  
  i = 0
  B <- diag(dim(A)[1])
  while(i < k){
    i <- i + 1
    B <- B %m% A
  }
  B
}
```

### (b)

```{r}
Z <- matrix(c(0.2, 0.6, 0.4, 0.7, 0.2, 0.1, 0.1, 0.2, 0.5), nrow = 3)
Z
Z %^% 0
Z %^% 5
Z %^% 50
Z %^% 500
```


## Question 5

```{r}
my_which <- function(X, arr.ind = FALSE){
  longvector <- as.vector(X)
  vec <- numeric(0)
  
  for(i in seq_len(length(longvector))){
    if(longvector[i] == TRUE){
      vec <- c(vec, i)
    }
  }
  
  vec
}



Z <- matrix(c(TRUE,FALSE,TRUE,FALSE,TRUE,FALSE,TRUE,TRUE,FALSE), nrow = 3)


my_which(Z)
which(Z)

# longvector <- as.vector(Z)
# longvector
# seq_len(longvector)
# which(Z)
# 
# 
# 
# Z
# which(Z)
# as.vector(Z)


```

## Question 6

```{r}
setwd("~/Desktop")
load("mlb.RData")

```

### (a)

### (i)
```{r}
tapply(hit, team, max)
```

### (ii)
```{r}
tapply(team, team, length)
table(team)
```

### (iii)
```{r}
tapply(hr, team, function(x){seq_along(which(x >= 1))[length(seq_along(which(x >= 1)))]})
```

### (b)

### (i)
```{r}
tapply((hit / ab)[ab > 100], team[ab > 100],max)
```

### (ii)

```{r}
tapply((hit / ab)[ab > 100], team[ab > 100], mean)
```

### (c)

```{r}
tapply(hr, list(pos, league), mean)
```

Right field had the largest difference between the two leagues.

### (d)

```{r}
# tapply(table(team, pos), table(team, pos), median)
# tapply()
# help(tapply)
# table(team, pos) 
table(pos) / length(levels(team))
# length(levels(team))
# table(team)
tapply(team, list(team, pos), length)
vapply(team, mean, numeric(1))
```

## Question 7 

### (a)

```{r}
my_row <- function(X){
  matrix(1:dim(X)[1], nrow = dim(X)[1], ncol = dim(X)[2])
}

X <- matrix(c(6, 3, 5, 2, 4, 1), nrow = 2)
Y <- matrix(c(1, 2, 3, 4, 5, 6, 7, 8, 9), nrow = 3)
row(X)
my_row(X)
row(Y)
my_row(Y)
```

### (b)

```{r}
my_col <- function(X){
  matrix(1:dim(X)[2], nrow = dim(X)[1], ncol = dim(X)[2], byrow = TRUE)
}

X <- matrix(c(6, 3, 5, 2, 4, 1), nrow = 2)
Y <- matrix(c(1, 2, 3, 4, 5, 6, 7, 8, 9), nrow = 3)
col(X)
my_col(X)
col(Y)
my_col(Y)
```

### (c)
```{r}
my_diag <- function(X, nrow, ncol){
  if(length(X) == 1){
    diagmat <- matrix(0 ,nrow = X, ncol = X)
    
    for(i in seq_len(dim(diagmat)[1])){
      diagmat[i,i] <- 1
    }
    
  }else if(length(dim(X)) > 1){
    diagmat <- numeric(0)
    longvector <- as.vector(X)

    for(i in seq_len(length(longvector))){
      if((sqrt(i) %% 1) == 0){
        diagmat <- c(diagmat, longvector[i])
      }
    }
  }else if(length(X) > 1){
    diagmat <- matrix(0, nrow = length(X), ncol = length(X))
    
    for(i in seq_len(length(X))){
      diagmat[i,i] <- X[i]
    }
    
  }
  diagmat
}

X <- matrix(c(6, 3, 5, 2, 4, 1), nrow = 2)

diag(5)
my_diag(5)
diag(c(1, 2, 3, 4, 5))
my_diag(c(1, 2, 3, 4, 5))
diag(X)
my_diag(X)

# length(dim(Y))
# diag(Y)
# mode(Y)
# diag(c(1))
# diag(X)
# typeof(Y)
# class(c(1))
# class(1)
# is.vector(5)
# diag(5)
# length(c(5, 2, 3, 4, 5))
# is.integer(4)
```


